---
title: "genoscapeRtools"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  github_document:
    toc: true
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "running-admixture-figs/"
)
```

## Introduction

The `genoscapeRtools` has a few functions that simplify the process of running 
the program `admixture` multiple times and multiple K values.  This is designed to
run on a Mac with a decent Unix-like system.  Hasn't been tested elsewhere.

## Data conversion

The first hurdle is to get your data into a proper PLINK "bed" file that `admixture` can
read.  This is pretty easy from the VCF file, but we are going to do it straight from the 
object that `read_012` creates. 

So, read that thing in:
```{r, read-the-file}
library(genoscapeRtools)

wifl_clean <- read_012("../cleaned_indv175_pos105000")
```
Then, if you want to make a plink bed file out of that you can do it like this:
```{r convert-to-bed}
convert_012_to_bed(wifl_clean, chromo_override = TRUE, prefix = "plinky")
```

That creates all these files:
```{r dir-it}
dir(pattern = "plinky.*")
```

And now you can pass the `plinky.bed` file to our function that runs admixture a lot of times.
That function will look for the .fam and the .bim files in the same place, so they have to be there
too.



## Running admixture with a single function.  
This function sets up a directory and runs admixture in parallel in all of them.   We use the defaults, which
puts all the results in a directory called "admixture_runs" in the current working directory.

We are going to just do one rep and only at K=3 and K=4
```{r run-admixture, eval=FALSE}
run_admixture(bed = "plinky.bed", Reps = 1, Kvals = 3:4)
```
The above can take a long time and it is hard to kill the process once you start it.  So, just let 'er go.

To see how things are coming along you can look at the files inside the `r_x_K_y` directories in the `admixture_runs`
directories where all the action is happening.


When it is all done, the output is in the directories like this:
```
2016-11-25 06:34 /tutorials/--% (master) ls admixture_runs/*
admixture_runs/data:
README.txt  input.bed   input.bim   input.fam

admixture_runs/r_1_K_3:
admixture.stderr  admixture.stdout  input.3.P         input.3.Q

admixture_runs/r_1_K_4:
admixture.stderr  admixture.stdout  input.4.P         input.4.Q

```

When that is done, read it into a data frame:
```{r slurp-it, message=FALSE}
admixQ <- slurp_admixture()
```

And then plot it quickly.
```{r plot-admix, width=14, height = 12}
ggplot_the_Qs(Qs = admixQ)
```
And, I need to add a cross-validation flag if we want to do cross-validation.


